### 2011-01-26 paper/learningMethods.tex
Updated HyperNEAT description

The first method is to choose a family of \emph{parametrized
  functions}.  By fixing the parameters to a particular set of values,
we obtain a deterministic motion function over time.  We tried several
parametrizations on the robot and, upon obtaining reasonable early
success, settled on one particular parametrization.  We dubbed this
the \emph{SineModel5}, as it has five parameters and employs a sine
wave as its root pattern.

@@

\[
\vec{g}(t) =
\left[ {\begin{array}{cc} 
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_F \cdot m_O & + offset \\ % 0
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_F           & + offset \\ % 1
a \cdot \sin(\frac{2\pi t}{\tau})                     & + offset \\ % 2
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_O           & + offset \\ % 3
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_R           & + offset \\ % 4
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_R \cdot m_O & + offset \\ % 5
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_F \cdot m_R & + offset \\ % 6
a \cdot \sin(\frac{2\pi t}{\tau}) \cdot m_F \cdot m_R & + offset \\ % 7
0                                                     & + offset \\ % 8
\end{array} } \right]
\]

@@

1) The value for $\vec{\theta}$ for the initial trial, and

2) The next value of $\vec{\theta}$ to try, given the $\vec{\theta}$
and $f(\vec{\theta})$ for all previous trials

@@

A CPPN is queried once for each link in the ANN substrate to determine that link's weight. The inputs to the CPPN are the Cartesian coordinates of both the source (e.g., $x = 2$, $y = 4$) and target (e.g., $x = 3$, $y = 5$) nodes of a link and a constant bias value. The CPPN takes these five values as inputs and produces one or two output values, depending on the substrate topology. If there is no hidden layer in the substrate, the single output is the weight of the link between a source node on the input layer and a target node on the output layer. If there is a hidden layer, the first output value determines the weight of the link between the associated input (source) and hidden layer (target) nodes, and the second output value determines the weight of the link between the associated hidden (source) and output (target) layer nodes. All pairwise combinations of source and target nodes are iteratively passed as inputs to a CPPN to deter-mine the weight of each substrate link. 
HyperNEAT is capable of exploiting the geometry of a problem: because the link values between nodes in the final ANN substrate are a function of the geometric positions of those nodes, HyperNEAT can exploit such information when solving a problem~\cite{stanley2009hypercube, clune2009sensitivity, clune2011performance}. In the case of quadruped locomotion, this property helped Hyper-NEAT produce gaits with front-back, left-right, and four-way symmetries~\cite{clune2009evolving, clune2011performance}.

>>>>

The first method is to choose a family of \emph{parametrized functions}.  By fixing the parameters to a particular set of values, we obtain a deterministic motion function over time.  We tried several parametrizations on the robot and, obtaining reasonble early success, settled on one particular parametrization.  We dubbed this the \emph{SineModel5}, as it has five parameters and employs a sine wave as its root pattern.

@@

\begin{verbatim}
        [g_0(t) ]    [ \amp * sin(\tau t) ]
        [g_1(t) ]    [ \amp * sin(\tau t) ]
        [g_2(t) ]    [ \amp * sin(\tau t) ]
g(t)   [g_3(t) ] = [ \amp * sin(\tau t) ]
        [g_4(t) ]    [ \amp * sin(\tau t) ]
        [g_5(t) ]    [ \amp * sin(\tau t) ]
        [g_6(t..) ]    [ \amp * sin(\tau t) ]
\end{verbatim}

@@

1) The value for $\vec{\theta}$ for the initial trial, and
2) The next value of $\vec{\theta}$ to try, given the $\vec{\theta}$ and $f(\vec{\theta})$ for all previous trials

@@

A CPPN is queried once for each link in the ANN substrate to determine that link's weight. The inputs to the CPPN are the Cartesian coordinates of both the source (e.g., $x = 2$, $y = 4$) and target (e.g., $x = 3$, $y = 5$) nodes of a link and a constant bias value. The CPPN takes these five values as inputs and produces one or two output values, depending on the substrate topology. If there is no hidden layer in the substrate, the single output is the weight of the link between a source node on the input layer and a target node on the output layer. If there is a hidden layer, the first output value determines the weight of the link between the associated input (source) and hidden layer (target) nodes, and the second output value determines the weight of the link between the associated hidden (source) and output (target) layer nodes. All pairwise combinations of source and target nodes are iteratively passed as inputs to a CPPN to determine the weight of each substrate link.

\edit{following paragraph is figure caption for file named: hyperneatExplanation.png}
HyperNEAT Produces ANNs from CPPNs. ANN weights are specified as a function of the geometric coordinates of the source node and the target node for each connection. The coordinates of these nodes and a constant bias are iteratively passed to the CPPN to determine each connection weight. The CPPN has two output values, which specify the weights for each connection layer as shown.

HyperNEAT is capable of exploiting the geometry of a problem: because the link values between nodes in the final ANN substrate are a function of the geometric positions of those nodes, HyperNEAT can exploit such information when solving a problem~\cite{stanley2009hypercube, clune2009sensitivity, clune2011performance}. In the case of quadruped locomotion, this property helped Hyper-NEAT produce gaits with front-back, left-right, and four-way symmetries~\cite{clune2009evolving, clune2011performance}.
